---
title: "Early Game Jungle Analysis"
subtitle: "Scuttle Dominance, Side Asymmetry, and Jungler Behavior"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    code-fold: true
    fig-width: 9
    fig-height: 5
execute:
  warning: false
  message: false
---

```{r setup}
library(supernova)
library(tidyverse)
library(supernova)
processed_data <- readRDS("processed_data.rds")
champion_profiles <- readRDS("champion_profiles.rds")
team_level_features <- readRDS("team_leveL_features.rds")
rawmatchdata <- readRDS("rawmatchdata.rds")
```

## Overview

This document explores early game jungle behavior in League of Legends using
match-level data. The central questions are:

- Does scuttle dominance translate into meaningful gold and pressure advantages?
- Do blue and red side differ in how they convert scuttle control?
- Does losing scuttles push junglers toward more aggressive, gank-oriented play?

---

## Dragon Takedown Timing

Before looking at scuttle dominance, it is worth establishing when dragon
contests are actually happening. The distributions below show first dragon
takedown time split by side, filtered to games where a dragon was taken.

```{r dragon-histogram}
processed_data %>%
  filter(earliestDragonTakedown > 0) %>%
  ggplot() +
  geom_histogram(aes(x = earliestDragonTakedown), color = "black", fill = "#c2dbff") +
  facet_wrap(~teamId) +
  theme_classic() +
  labs(
    title    = "First Dragon Takedown Time",
    subtitle = "Split by Blue (100) and Red (200) teams",
    x        = "Earliest Dragon Takedown (seconds)",
    y        = "Count"
  )
```

```{r dragon-boxplot}
processed_data %>%
  filter(earliestDragonTakedown > 0) %>%
  ggplot() +
  geom_boxplot(aes(y = earliestDragonTakedown), fill = "#c2dbff") +
  facet_grid(~teamId) +
  theme_classic() +
  labs(
    title    = "First Dragon Takedown Time — Spread",
    subtitle = "Split by Blue (100) and Red (200) teams",
    y        = "Earliest Dragon Takedown (seconds)"
  )
```

---

## Champion Sample Sizes by Role

To contextualize the champion-level analyses, the chart below shows average
games played per champion broken out by role. Roles with fewer games per
champion may produce noisier estimates downstream.

```{r games-by-role}
champion_profiles %>%
  group_by(teamPosition) %>%
  summarise(avg_games = mean(games_played, na.rm = TRUE)) %>%
  ggplot(aes(x = teamPosition, y = avg_games, fill = teamPosition)) +
  geom_col() +
  theme_classic() +
  labs(
    title = "Average Games Played per Champion by Role",
    x     = NULL,
    y     = "Average Games Played",
    fill  = NULL
  )
```

---

## Team and Rift Herald

A quick correlation check between Red vs. Blue team and if they took the Rift Herald shows that blue team takes 

```{r cs-herald-cor}
cor.test(processed_data$teamId, processed_data$teamRiftHeraldKills)
processed_data %>%
  group_by(teamId) %>%
  count(teamRiftHeraldKills) %>%
  mutate(
    teamId = if_else(teamId == 100, "Blue", "Red"),
  ) %>%
  pivot_wider(names_from = teamRiftHeraldKills, values_from = n)
```

---

## Gold by Scuttle Dominance and Side

Here we move to the central question: does scuttle dominance translate into
gold leads, and does blue or red side convert it more efficiently?

Each team's `initialCrabCount` reflects their own scuttle haul (0, 1, or 2),
making this a true per-team measure rather than a derived gap.

```{r gold-anova}
gold_cols <- c("top_gold", "jungle_gold", "mid_gold", "adc_gold", "supp_gold")

team_analysis <- team_level_features %>%
  mutate(
    side = if_else(teamId == 100, "Blue", "Red"),
    crab_group = factor(initialCrabCount,
                        levels = c(0, 1, 2),
                        labels = c("No Scuttle", "One Scuttle", "Full Control"))
  )

gold_anova_results <- map_df(gold_cols, function(col) {
  formula <- as.formula(paste(col, "~ crab_group * side"))
  fit     <- lm(formula, data = team_analysis)
  f       <- summary(fit)$fstatistic

  tibble(
    feature    = col,
    F_stat     = round(f[1], 3),
    p_value    = pf(f[1], f[2], f[3], lower.tail = FALSE),
  )
}) %>%
  arrange(p_value) %>%
  mutate(
    p_adjusted  = p.adjust(p_value, method = "BH"),
    significant = p_adjusted < 0.05
  )

gold_anova_results
```

```{r gold-plot}
team_analysis %>%
  select(crab_group, side, all_of(gold_cols)) %>%
  pivot_longer(cols = all_of(gold_cols),
               names_to  = "metric",
               values_to = "gold") %>%
  group_by(crab_group, side, metric) %>%
  summarise(mean_gold = mean(gold, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = crab_group, y = mean_gold, color = side, group = side)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Blue" = "steelblue", "Red" = "tomato")) +
  facet_wrap(~ metric, scales = "free_y") +
  theme_minimal() +
  labs(
    title    = "Mean Gold by Scuttle Dominance and Side",
    subtitle = "Does blue or red side convert scuttle dominance into gold more efficiently?",
    x        = "Scuttle Dominance",
    y        = "Mean Gold",
    color    = NULL
  )
```

---

## Jungler Kill Participation and Proximity

If junglers losing scuttles are pushed out of the river, a hypothesis is that they will instead spend
their time lane ganking, causing proximity percent to go up and KP to increase compared to double crabs. 

```{r participation-setup}

jungle_involvement <- rawmatchdata %>%
  filter(teamPosition == "JUNGLE") %>%
  mutate(kill_participation = kills + assists) %>%
  select(matchId, teamId, kill_participation)

team_analysis <- team_level_features %>%
  left_join(jungle_involvement, by = c("matchId", "teamId")) %>%
  mutate(
    side = if_else(teamId == 100, "Blue", "Red"),
    crab_group = factor(initialCrabCount,
                        levels = c(0, 1, 2),
                        labels = c("No Scuttle", "One Scuttle", "Full Control"))
  )
```

```{r participation-anova}
fit <- lm(kill_participation.x ~ crab_group * side, data = team_analysis)
summary(fit)
```

```{r participation-plot}
prox_cols <- c("top_prox", "mid_prox", "adc_prox", "supp_prox")

team_analysis %>%
  select(crab_group, side, kill_participation.x, all_of(prox_cols)) %>%
  pivot_longer(cols = -c(crab_group, side),
               names_to  = "metric",
               values_to = "value") %>%
  group_by(crab_group, side, metric) %>%
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = crab_group, y = mean_value, color = side, group = side)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Blue" = "steelblue", "Red" = "tomato")) +
  facet_wrap(~ metric, scales = "free_y") +
  theme_minimal() +
  labs(
    title    = "Jungler Kill Participation and Proximity by Scuttle Dominance",
    subtitle = "Tests whether losing scuttles leads to more aggressive/gank-oriented play",
    x        = "Scuttle Dominance",
    y        = "Mean Value",
    color    = NULL
  )
```

---

## Key Takeaways

- **Red side consistently earns more gold** across all roles regardless of
  scuttle count, though the rate of increase with scuttle dominance is similar
  for both sides. The exception is top lane, where red side pulls ahead more
  sharply at Full Control.

- **Kill participation rises with scuttle dominance** rather than falling,
meaning that of course junglers who win both scuttles will also have more 
agency on the entire map. 

- **Lane proximity drops across all roles as scuttle dominance increases**,
  even though kill participation is rising. This points to scuttle-dominant
  junglers collecting kills through river skirmishes rather than traditional
  ganks.

- **The working heuristic:** a jungler losing scuttles is more likely to seek
  kill participation through lane ganks; a jungler winning scuttles is
  collecting it through river and objective fights.
